\input{header.tex}

\title{Solutions Sheet 7}
\author{Yannis B\"{a}hni}
\address[Yannis B\"{a}hni]{University of Zurich, R\"{a}mistrasse 71, 8006 Zurich}
\email[Yannis B\"{a}hni]{\href{mailto:yannis.baehni@uzh.ch}{\nolinkurl{yannis.baehni@uzh.ch}}}

\begin{document}

\maketitle
\thispagestyle{fancy}

\setcounter{section}{1}

\begin{enumerate}[label = \textbf{Exercise \arabic*.},wide = 0pt, itemsep = 1.5ex]
	\item
	\item
	\item 
		\begin{lemma}
			Let $y \in H$ and define a mapping $\varphi_y : H \to \Cbb$ by $\varphi_y(x) := \langle A(y),x \rangle$. Then $\varphi_y \in \Lcal(H,\Cbb)$.
			\label{lem:family}
		\end{lemma}

		\begin{proof}
			Clearly, $\varphi_y$ is linear since $\langle \cdot,\cdot \rangle$ is linear in the second component. Moreover, $\varphi_y$ is bounded. Indeed, using Cauchy-Schwarz yields
			\begin{equation*}
				\abs[0]{\varphi_y(x)} = \abs[0]{\langle A(y), x \rangle} \leq \norm{A(y)} \norm{x}
			\end{equation*}
			\noindent for all $x \in H$.
		\end{proof}

		Thus we may define a family
		\begin{equation*}
			\Fcal := \cbr{\varphi_y : y \in \partial B_1(0)} \subseteq \Lcal(H,\Cbb).
		\end{equation*}

		Let $x \in H$. Then for any $y \in \partial B_1(0)$ we have that
		\begin{equation*}
			\abs[0]{\varphi_y(x)} = \abs[0]{\langle A(y),x \rangle} = \abs[0]{\langle y,A(x) \rangle} \leq \norm{y} \norm{A(x)} = \norm{A(x)}
		\end{equation*}
		\noindent by symmetry and again Cauchy-Schwarz. Hence
		\begin{equation*}
			\sup_{T \in \Fcal}\abs[0]{T(x)} = \sup_{y \in \partial B_1(0)} \abs[0]{\varphi_y(x)} \leq \norm{A (x)}
		\end{equation*}
		\noindent for all $x \in H$. Since any Hilbert space is a Banach space, an application of \emph{Banach-Steinhaus} yields the existence of a constant $c > 0$ such that
		\begin{equation*}
			\sup_{T \in \Fcal}\norm[0]{T} = \sup_{y \in \partial B_1(0)} \norm[0]{\varphi_y} \leq c.
		\end{equation*}
		For $x \in H$ such that $A(x) \neq 0$ we have that
		\begin{align*}
			\norm{A(x)}^2 &= \langle A(x),A(x) \rangle\\
			&= \norm{x} \langle A(x / \norm{x}), A(x)\rangle\\
			&= \norm{x} \varphi_{x/\norm{x}}(A(x))\\
			&\leq \norm{x} \abs[0]{\varphi_{x/\norm{x}}(A(x))}\\
			&\leq \norm{x} \norm{A(x)} \norm[0]{\varphi_{x/\norm{x}}}\\
			&\leq c\norm{x} \norm{A(x)}
		\end{align*}
		\noindent and thus dividing both sides by $\norm{A(x)}$ yields the boundedness of $A$.

	\item

	\item
		~
		\begin{enumerate}[label = \textbf{\alph*.},wide = 0pt, itemsep = 1.5ex]
			\item We define
				\begin{equation*}
					\Fcal := \cbr{B(\cdot,y) : y \in \partial B_1(0)}.
				\end{equation*}

				\begin{lemma}
					We have that $\Fcal \subseteq \Lcal(X,\Kbb)$ and for all $x \in X$, there exists $c_x \geq 0$ such that $\sup_{T \in \Fcal}\abs{T(x)} \leq c_x$.
				\end{lemma}

				\begin{proof}
					Let $y \in \partial B_1(0)$. Then $B(\cdot,y)$ is linear by definition of a bilinear functional. Moreover, for any $x \in X$ we have that
					\begin{equation*}
						\abs{B(x,y)} \leq c_y \norm{x}
					\end{equation*}
					\noindent for some $c_y \geq 0$ by continuity of $B$ in the first argument. Hence $\Fcal \subseteq \Lcal(X,\Kbb)$. Let $x \in X$. Then
					\begin{equation*}
						\abs{B(x,y)} \leq c_x \norm{y} = c_x
					\end{equation*}
					\noindent for some $c_x \geq 0$ by continuity of $B$ in the second argument. Thus
					\begin{equation*}
						\sup_{T \in \Fcal}\abs{T(x)} = \sup_{y \in \partial B_1(0)} \abs{B(x,y)} \leq c_x
					\end{equation*}
					\noindent for all $x \in X$.
				\end{proof}

				An application of \emph{Banach-Steinhaus} on the family $\Fcal$ yields the existence of a constant $c \geq 0$ such that
				\begin{equation*}
					\sup_{T \in \Fcal}\norm{T} \leq c.
				\end{equation*}
				Let $x,y \in X$. Then
				\begin{align*}
					\abs{B(x,y)} &= \norm{x}\norm{y} \abs[0]{B(x/\norm{x},y/\norm{y})}\\
					&\leq \norm{x}\norm{y} \sup_{\norm[0]{\xi} = 1}\abs[0]{B(\xi,y/\norm{y})}\\
					&\leq \norm{x}\norm{y} \sup_{\norm[0]{\zeta} = 1} \sup_{\norm[0]{\xi} = 1} \abs{B(\xi,\zeta)}\\
					&= \norm{x}\norm{y} \sup_{\norm[0]{\zeta} = 1}\norm{B(\cdot,\zeta)}\\
					&\leq c\norm{x}\norm{y}.
				\end{align*}

				\begin{lemma}
					Equip $X \times X$ with the norm $\norm{(x,y)} := \norm{x} + \norm{y}$. Then $B$ is continuous.
				\end{lemma}

				\begin{proof}
					Let $(x,y) \in X \times X$ and $(x_n,y_n)_{n \in \Nbb}$ be a sequence in $X \times X$ converging to $(x,y)$. We claim that $x_n \to x$ and $y_n \to y$ in $X$. Indeed
					\begin{equation*}
						\norm{x_n - x} \leq \norm{x_n - x} + \norm{y_n - y} = \norm{(x_n,y_n) - (x,y)} \to 0
					\end{equation*}
					\noindent as $n \to \infty$ and similarly
					\begin{equation*}
						\norm{y_n - y} \leq \norm{x_n - x} + \norm{y_n - y} = \norm{(x_n,y_n) - (x,y)} \to 0.
					\end{equation*}
					Moreover, since $y_n \to y$, $y_n$ is bounded, i.e. there exists some $M \geq 0$ such that $\norm{y_n} \leq M$ for all $n \in \Nbb$. Hence
					\begin{align*}
						\abs{B(x_n,y_n) - B(x,y)} &= \abs{B(x_n,y_n) - B(x,y_n) + B(x,y_n) - B(x,y)}\\
						&= \abs{B(x_n - x,y_n) + B(x,y_n - y)}\\
						&\leq \abs{B(x_n - x,y_n)} + \abs{B(x,y_n - y)}\\
						&\leq c \norm{x_n - x}\norm{y_n} + c \norm{x}\norm{y_n - y}\\
						&\leq cM \norm{x_n - x} + c\norm{x}\norm{y_n - y} \rightarrow 0
					\end{align*}
					\noindent as $n \to \infty$.
				\end{proof}
			\item
				~
				\begin{lemma}
					$B$ is a bilinear functional on $\cal{P}$ which is continuous in each argument separately.
				\end{lemma}

				\begin{proof}
					The bilinearity of $B$ directly follows from the linearity of the integral. Fix $q \in \cal{P}$. Then for any $p \in \cal{P}$ we have that
					\begin{equation*}
						\abs{B(p,q)} = \abs[3]{\int_0^1 p(t)q(t)dt} \leq \int_0^1 \abs{p(t)}\abs{q(t)}dt \leq \sup_{t \in \intcc{0,1}}\abs{q(t)} \int_0^1 \abs{p(t)}dt = c_q \norm{p}
					\end{equation*}
					\noindent since $q$ is continuous. Similarly, for each fixed $p \in \cal{P}$ we get that $\abs{B(p,q)} \leq c_p \norm{q}$ for all $q \in \cal{P}$.
				\end{proof}
		\end{enumerate}
\end{enumerate}

\end{document}
