\input{../latex/header.tex}

\title{Solutions Sheet 1}
\author{Yannis B\"{a}hni}
\address[Yannis B\"{a}hni]{University of Zurich, R\"{a}mistrasse 71, 8006 Zurich}
\email[Yannis B\"{a}hni]{\href{mailto:yannis.baehni@uzh.ch}{yannis.baehni@uzh.ch}}

\begin{document}
\maketitle
\thispagestyle{fancy}

\begin{enumerate}[label = \textbf{Exercise \arabic*.},wide = 0pt, itemsep=1.5ex]
\item ~ 
	\begin{enumerate}[label = \textbf{\alph*.},wide = 0pt, itemsep=1.5ex]
		\item The first part can be shown for an arbitrary set $X$. Clearly $\varnothing,X \in \Tcal$ since $X^c = \varnothing$ is countable. Let $(U_\iota)_{\iota \in I}$ be a family of sets in $\Tcal$. If $U_\iota = \varnothing$ for all $\iota \in I$ we have that $\cup_{\iota \in I} U_\iota = \varnothing \in \Tcal$. So assume that $U_{\iota_0} \neq \varnothing$ for some $\iota_0 \in I$. But then $U^c_{\iota_0}$ is countable, and so is $(\cup_{\iota \in I} U_\iota)^c = \cap_{\iota \in I}U_\alpha^c \subseteq U^c_{\iota_0}$. Lastly, let $U_1,\dots,U_n \in \Tcal$ for $n\in \mathbb{Z}$, $n \geq 1$. If $U_\iota = \varnothing$ for some $\iota$, then $\cap_{\iota = 1}^n U_\iota = \varnothing$ and thus $\cap_{\iota = 1}^n U_\iota \in \Tcal$. So assume that $U_\iota \neq \varnothing$ for $\iota = 1,\dots,n$. Then $(\cap_{\iota = 1}^n U_\iota)^c = \cup_{\iota = 1}^n U_\iota^c$ which is a finite union of countable sets, which is countable. Hence $\Tcal$ is indeed a topology on $X$.\\
			We claim that $(X,\Tcal)$ is not Hausdorff when $X$ is uncountable. Towards a contradiction assume that $(X,\Tcal)$ is Hausdorff. Let $p,q \in X$ with $p \neq q$. Hence there exist (open) neighbourhoods $U$ and $V$ of $p$ and $q$ respectively such that $U \cap V = \varnothing$. Now $X = U \cup U^c$, where $U^c$ is countable and clearly nonempty. But $U \cap V = \varnothing$ implies $U \subseteq V^c$ which therefore yields that $U$ is also countable. Hence $X$ is a union of two countable sets and thus countable. Contradiction.
		\item We prove both times the contrapositive. Assume that there is a family $(A_\iota)_{\iota \in I}$ of closed subsets of $X$ having the finite intersection property such that $\cap_{\iota \in I} A_\iota = \varnothing$. Then $\cup_{\iota \in I} A^c_\iota = (\cap_{\iota \in I} A_\iota)^c = X$. Since each $A_\iota$ is closed, $A_\iota^c$ is open for all $\iota \in I$ and thus $(A^c_\iota)_{\iota \in I}$ is an open cover for $X$. We claim that $(A_\iota^c)_{\iota \in I}$ does not admit any finite subcover. Towards a contradiction, assume that it does. Hence we find $\iota_1,\dots,\iota_n \in I$, $n \in \Zbb$, $n \geq 1$, such that $\cup_{k = 1}^n A_{\iota_k}^c = X$. But then $\cap_{k = 1}^n A_{\iota_k} = \varnothing$, contradicting the finite intersection property of the family $(A_\iota)_{\iota \in I}$.\\
			Conversly, suppose that there exists an open cover $(A_\iota)_{\iota \in I}$ of $X$ which does not admit a finite subcover. We claim that the closed family $(A_\iota^c)_{\iota \in I}$ has the finite intersection property and $\cap_{\iota \in I}A_\iota = \varnothing$. Let $\iota_1,\dots,\iota_n \in I$, $n \in \Zbb$, $n \geq 1$. Since $(A_{\iota_k})_{k = 1}^n$ cannot cover $X$, otherwise it would be a finite subcover of $(A_\iota)_{\iota \in I}$, we have that $\cap_{k = 1}^n A_{\iota_k}^c \neq \varnothing$. Thus $(A_\iota^c)_{\iota \in I}$ has the finite intersection property. Since $(A_\iota)_{\iota \in I}$ covers $X$ we have that $\cap_{\iota \in I} A_\iota^c = \varnothing$.
	\end{enumerate}
\item
	~
	\begin{enumerate}[label = \textbf{\alph*.},wide = 0pt, itemsep=1.5ex]
		\item Clearly, $\varnothing,X \in \Tcal_d$. Let $(U_\iota)_{\iota \in I}$ be a family of elements in $\Tcal_d$ and $x \in \cup_{\iota \in I}U_\iota$. Then there exists $\iota \in I$ such that $x \in U_\iota$. Furthermore, we find $\varepsilon > 0$ such that $B_\varepsilon(x) \subseteq U_\iota$. Hence $B_\varepsilon(x) \subseteq \cup_{\iota \in I} U_\iota$. Let $U_1,\dots,U_n \in \Tcal$ for $n\in \mathbb{Z}$, $n \geq 1$, and $x \in \cap_{\iota = 1}^n U_\iota$. Hence there exist $\varepsilon_1,\dots,\varepsilon_n > 0$ such that $B_{\varepsilon_\iota}(x) \subseteq U_\iota$ for $\iota = 1,\dots,n$ and so $B_{\wtilde{\varepsilon}}(x) \subseteq \cap_{\iota = 1}^n U_\iota$ for $\wtilde{\varepsilon} := \min\cbr[0]{\varepsilon_1,\dots,\varepsilon_n}$. Thus $\Tcal_d$ is a topology on $X$. 
		\item We will use the fact that two metrics induce the same topology if and only if they induce the same convergence. Let $\wtilde{M} := \intoo{0,\infty}$. Define $f : \wtilde{M} \to \wtilde{M}$ by $f(x) := 1/x$. Then clearly $d_2 = \wtilde{d_2}\vert_M$ and $d_1 = \wtilde{d_1}\vert_M$, where 
		\begin{equation*}
			\wtilde{d_2} : \wtilde{M} \times \wtilde{M} \xrightarrow{f \times f}{} \wtilde{M} \times \wtilde{M} \xrightarrow{\abs[0]{\cdot,\cdot}}{} \Rbb
		\end{equation*}
		\noindent and
		\begin{equation*}
			\wtilde{d_1} : \wtilde{M} \times \wtilde{M} \xrightarrow{f \times f}{} \wtilde{M} \times \wtilde{M} \xrightarrow{\wtilde{d_2}}{} \Rbb.
		\end{equation*}
		It is easy to show that $\wtilde{d_2}$ is a metric. Let $x \in M$ and $(x_n)_{n \in \mathbb{N}}$ be a sequence in $M$. Assume that $x_n \xrightarrow{d_1}{} x$. Then 
		\begin{equation*}
			d_2(x_n,x) = \wtilde{d_1}\del[1]{f(x_n),f(x)} \to 0
		\end{equation*}
		\noindent and
		\begin{equation*}
			d_1(x_n,x) = \wtilde{d_2}\del[1]{f(x_n),f(x)} \to  0
		\end{equation*}
		\noindent by the continuity of $f$ on $\wtilde{M}$.\\
		$(M,d_1)$ is complete since $M$ is a closed subset of the complete metric space $\Rbb$. Consider the sequence $(n)_{n \in \mathbb{N}}$ in $M$. Clearly, it is a Cauchy sequence in $(M,d_2)$ since $\frac{1}{n} \xrightarrow{\abs[0]{\cdot}}{} 0$. Assume that it converges also in $(M,d_2)$. Since the induced topologies of $d_1$ and $d_2$ are the same, we would get that $(n)_{n \in \Nbb}$ also converges in $(M,d_1)$. But this is absurd. Hence $(M,d_2)$ cannot be complete. 
	\end{enumerate}
\end{enumerate}
\printbibliography
\end{document}
